Tree establishment in models of forest dynamics: a key priority for further research

### 2023-03-09 manuscript/manuscriptRW.Rmd
add changes till conclussions

We recommend that the further development of the establishment modules
is implemented as additional features that can be traced back, as done
in this study for the model variants of ForClim and ForCEEPS, and that
model complexity and structure must always be connected with model
objectives [@albrich2020]. If it should be necessary to include more
detail in the establishment models, this will come with higher
parameterisation efforts, which will most likely lead to lower
generalization as the required data will have to be collected from
specific locations, as there is no general establishment dataset
available.

@@

Here we have only consider tree establishment in the equilibrium state.
It is equally important to understand how these models project tree
establishment after changes in forest structure like those brought about
by disturbances [@seidl2022], or under different management strategies
[@lindner2000]. However, this will require an entirely different set of
observed data, and potentially not all models of forest dynamics would
be able to assess the relationship of these aspects on tree
establishment, e.g., due to the lack of disturbance or appropriate
management modules.

We recommend to further investigate the implications of the current
modeling strategies for tree establishment, particularly those related
to excess tree establishment and probably late excess mortality, by
focusing on a wider range of tree sizes and the related ingrowth
dynamics. We also recommend, especially for the landscape-level models,
the inclusion of explicit spatial considerations regarding tree ingrowth
[@beckage2003]; this, however, is facing a serious challenge regarding
appropriate datasets, as inventory data have wide coverage, but do not
allow for the assessment of spatial interaction effects.

@@

Models of forest dynamics are important tools in science and decision
support, and how tree establishment is defined has strong implications
for simulated forest properties. The 15 models and variants used here
are facing similar challenges when representing tree establishment. They
generally overestimate tree establishment rates, and the simulated
establishment niche is not always captured accurately as a function of
the biotic (light) and abiotic (temperature and moisture).

@@

Regarding mortality in the early phase of tree life, many models that
feature a particularly high overestimation of ingrowth levels are
compensating this by a larger tree mortality. Often, however, this
compensation is not sufficient to reduce the high ingrowth levels to
realistic values. Overall, however, there is no clear pattern across all
models.

When capturing tree establishment, the specific design decisions taken
in the development of any model are more important for its behavior
(accuracy) than scale (stand, landscape, global), background (empirical
vs. process-based), and complexity. Having both empirical and
process-based models in our set, the empirically-based models could have
been expected to have a better performance, as they were calibrated with
inventory data, but this was not the case. Similarly, higher model
complexity does not represent an improvement for capturing tree
establishment.

>>>>

We recommend that the further development of the establishment modules
is implemented as additional features that can be traced back, as done
in this study for the model variants of ForClim and ForCEEPS, and that
model complexity and structure must always be connected with model
objectives [@albrich2020]. If it should be necessary to include more
detail in the establishment models, this will come with higher
parameterisation efforts, which will most likely lead to lower
generalization as the required data will have to be collected from
specific locations, as currently there is no general establishment
dataset available.

@@

Here we have only considered tree establishment in the equilibrium
state. It is equally important to understand how these models project
tree establishment after changes in forest structure by disturbances
[@seidl2022], or under different management strategies [@lindner2000].
However, this will require an entirely different set of observed data,
and potentially not all models of forest dynamics would be able to
assess the relationship of these aspects on tree establishment, e.g.,
due to the lack of disturbance or appropriate management modules.

We recommend to further investigate the implications of the current
modeling strategies for tree establishment, particularly those related
to excess tree establishment and late excess mortality, by focusing on a
wider range of tree sizes and the related ingrowth dynamics. We also
recommend, especially for the landscape-level models, the inclusion of
explicit spatial considerations regarding tree ingrowth [@beckage2003];
this, however, is facing a serious challenge regarding appropriate
datasets, as inventory data have wide coverage, but do not allow for the
assessment of spatial interaction effects.

@@

Models of forest dynamics are important tools in science and decision
support, and how tree establishment is defined has strong implications
for simulated forest properties. The 15 models and variants used here
are facing similar challenges when representing tree establishment. They
generally overestimate tree establishment rates, and the simulated
establishment niche is not always captured accurately as a function of
the biotic (light) and abiotic (temperature and moisture) factors.

@@

Regarding mortality in the early phase of tree life, many models that
feature a particularly high overestimation of ingrowth levels are
compensating for this by a larger tree mortality. Often, this
compensation is not sufficient to reduce the high ingrowth levels to
realistic values. Overall, there is no clear mortality pattern across
all models.

When capturing tree establishment, the specific design decisions taken
in the development of any model are more important for its behavior
(accuracy) than scale (stand, landscape, global), modelling approach
(empirical vs. process-based), and complexity. Having both empirical and
process-based models in our set, the empirically-based models could have
been expected to have a better performance, as they were calibrated with
inventory data, but this was not the case. Similarly, higher model
complexity does not represent an improvement for capturing tree
establishment.
